---
title: "HMRC Euro to GBP Monthly Rates 2025/26"
format: html
editor: visual
output-file: index.html
output-dir: docs
---

```{r setup, include=FALSE}
library(tidyverse)
library(httr2)
library(jsonlite)
library(lubridate)
library(kableExtra)
```

## Monthly rate extraction

This code retrieves the monthly EUR→GBP exchange rates for the 2025-26 financial year by calling the open HMRC exchange‑rate API developed by [Matías J. S](https://github.com/matchilling/hmrc-exchange-rates). and assembling the Euro rate for each month into a single tidy table.

You can double check values in the official site [HMRC monthly Euro to Sterling exchange rates](https://www.trade-tariff.service.gov.uk/exchange_rates)

```{r fetch-rates-final}
# Define months for FY 2025/26 (April 2025 - March 2026)
fy_months <- seq(as.Date("2025-04-01"), as.Date("2026-03-01"), by = "month") %>%
  format("%Y-%m")

# Function to fetch rate for a month
fetch_hmrc_rate <- function(month_ym) {
  year_mon <- str_replace(month_ym, "-", "/")
  url <- paste0("https://hmrc.matchilling.com/rate/", year_mon, ".json")
  
  resp <- request(url) %>%
    req_timeout(10) %>%
    req_perform()
  
  if (resp_status(resp) == 200) {
    data <- resp %>%
      resp_body_json(simplifyVector = TRUE) %>%
      .$rates %>%
      enframe(name = "currency", value = "rate") %>%
      filter(currency == "EUR") %>%
      mutate(
        month = month_ym,
        period_start = ymd(paste0(month_ym, "-01")),
        period_end = period_start %m+% months(1) - days(1),
        eur_per_gbp = as.numeric(rate)
      ) %>%
      select(month, eur_per_gbp, period_start, period_end)
    return(data)
  } else {
    tibble(
      month = month_ym, 
      eur_per_gbp = NA_real_,
      period_start = ymd(paste0(month_ym, "-01")),
      period_end = period_start %m+% months(1) - days(1)
    )
  }
}

# Fetch all months
rates <- map_df(fy_months, fetch_hmrc_rate)

```

## View the table

This bit creates a simple table visualization for the html file

```{r summary-table-kable}
rates %>%
  mutate(
    month_display = month,
    eur_per_gbp = round(eur_per_gbp, 4),
    gbp_per_eur = round(1/eur_per_gbp, 4)
  ) %>%
  select(month_display, eur_per_gbp, gbp_per_eur) %>%
  kbl(
    caption = "HMRC Monthly EUR/GBP Rates FY 2025/26",
    col.names = c("Month", "€ per £1", "£ per €1")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  # Fixed: pack_rows() doesn't accept font_size argument
  pack_rows("2025", 1, 9) %>%
  pack_rows("2026", 10, 12)

```

## Export it for your Tax Returns

Here, we save the csv file with the 12 months you need and the currency you need. (if you are looking for any other currency , just change the line in the first chunck of code to match your currency **filter(currency == "EUR")**

```{r export}
# Save clean CSV for HMRC filing
rates_export <- rates %>%
  select(month, eur_per_gbp) %>%
  mutate(
    eur_per_gbp = round(eur_per_gbp, 4),
    gbp_per_eur = round(1/eur_per_gbp, 4)
  )

write_csv(rates_export, "hmrc_eur_gbp_2025_26.csv")
print("CSV saved: hmrc_eur_gbp_2025_26.csv")

# Preview
rates_export
```
