---
title: "HMRC Euro to GBP Monthly Rates 2025/26"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tidyverse)
library(httr2)
library(jsonlite)
library(lubridate)
library(kableExtra)
```

If you only need a list with the monthly Euro to Sterling rates for the financial year 2025-26:

## Monthly rate extraction

This code retrieves monthly EUR/GBP rates using the open HMRC Exchange Rates API created by [Matías J. Schilling](https://github.com/matchilling/hmrc-exchange-rates)

```{r fetch-rates-final}
# Define months for FY 2025/26 (April 2025 - March 2026)
fy_months <- seq(as.Date("2025-04-01"), as.Date("2026-03-01"), by = "month") %>%
  format("%Y-%m")

# Function to fetch rate for a month
fetch_hmrc_rate <- function(month_ym) {
  year_mon <- str_replace(month_ym, "-", "/")
  url <- paste0("https://hmrc.matchilling.com/rate/", year_mon, ".json")
  
  resp <- request(url) %>%
    req_timeout(10) %>%
    req_perform()
  
  if (resp_status(resp) == 200) {
    data <- resp %>%
      resp_body_json(simplifyVector = TRUE) %>%
      .$rates %>%
      enframe(name = "currency", value = "rate") %>%
      filter(currency == "EUR") %>%
      mutate(
        month = month_ym,
        period_start = ymd(paste0(month_ym, "-01")),
        period_end = period_start %m+% months(1) - days(1),
        eur_per_gbp = as.numeric(rate)
      ) %>%
      select(month, eur_per_gbp, period_start, period_end)
    return(data)
  } else {
    tibble(
      month = month_ym, 
      eur_per_gbp = NA_real_,
      period_start = ymd(paste0(month_ym, "-01")),
      period_end = period_start %m+% months(1) - days(1)
    )
  }
}

# Fetch all months
rates <- map_df(fy_months, fetch_hmrc_rate)

```

## View the table

```{r summary-table-kable}
rates %>%
  mutate(
    month_display = month,
    eur_per_gbp = round(eur_per_gbp, 4),
    gbp_per_eur = round(1/eur_per_gbp, 4)
  ) %>%
  select(month_display, eur_per_gbp, gbp_per_eur) %>%
  kbl(
    caption = "HMRC Monthly EUR/GBP Rates FY 2025/26",
    col.names = c("Month", "EUR per £1", "£1 per EUR")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  # Fixed: pack_rows() doesn't accept font_size argument
  pack_rows("2025", 1, 9) %>%
  pack_rows("2026", 10, 12)
```

## Export it for your Tax Returns

In a simple csv file with only the 12 month you need and the currency you need. (if you are looking for any other currency , just change the bit in the code **filter(currency == "EUR")**

```{r export}
# Save clean CSV for HMRC filing
rates_export <- rates %>%
  select(month, eur_per_gbp) %>%
  mutate(
    eur_per_gbp = round(eur_per_gbp, 4),
    gbp_per_eur = round(1/eur_per_gbp, 4)
  )

write_csv(rates_export, "hmrc_eur_gbp_2025_26.csv")
print("CSV saved: hmrc_eur_gbp_2025_26.csv")

# Preview
rates_export
```
